\documentclass[../OpenAppliedMusicTheory.tex]{subfiles}

\externaldocument{../OpenAppliedMusicTheory}

\begin{document}
    
    \sfchapter{Meta}{0}
    
    While there is a good deal of information in the about section (\pref{}{iii}), there are a few essentials that I would say are important for "everyone" to read. I know many will probably skip around between chapters and not read this, but that's okay. It totally makes sense to skip stuff you know well and if that's the best way to make use of this book, I encourage it! Anyways, moving on...
    
    \section{References}\label{ch0.references}

    There are often times places where the book will mention an idea or topic that has been discussed in more detail before (lower chapter numbers is all this means), or will be discussed in more detail in future chapters. It's understandable that our teaching might not be the best or that certain things might be easy to forget or hard to remember the details of. Additionally, we understand if you want to skip around in the book. To make navigation easier, we will include references that look like \pref{98.76.5}{31415}. The first number (98) represents the chapter number, the second number (76) represents the section number, and the third number (5) represents the subsection number. Additionally, this is just the number that would be seen before a chapter/section/subsection %is this confusing because of the slashes
     title. The first two numbers are what would be displayed in the table of contents for a chapter or section. The number after the slash (31415) is the page number where the topic appears. We will scatter these references throughout the book.

\end{document}